---
title: "Winning the Lottery (Powerpoint)"
author: "Mario Arce Acosta"
date: "2024-06-14"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3


```{r H-1B_Data}
library(readxl)
library(readr)
library(ggplot2)
library(forecast)
library(stargazer)
library(kableExtra)

#USER XCITE: laptop"C:\Users\XCITE-admin\Documents\GitHub\Regular_R_Work\H1B_data.xlsx"
#USER mario:
H1B_Data <- read_csv("C:/Users/mario/Documents/GitHub/Regular_R_Work/H1B_Data.csv")
View(H1B_Data)

LPR_data <- read_excel("C:/Users/mario/Documents/GitHub/Regular_R_Work/LPR_1820-2022_data.xlsx", sheet = "LPR obtainments")
View(LPR_data)

# fearData<- read_excel("C:/Users/mario/Desktop/Datasets/Migration_Fear_EPU_Data.xlsx", sheet = "Sheet1")
# View(fearData)

# removalsData <- read_excel("C:/Users/mario/Desktop/Datasets/Alien_removal_returns_fy2019_data.xlsx", sheet = "(R) Aliens removed or returned")
# View(removalsData)
```

```{r  H1B_Data Clean & Prepare}
##Work began at XCITE

#Name Columns
colnames(H1B_Data)[1]<-"Year"
colnames(H1B_Data)[8]<-"Initial_Approval"
colnames(H1B_Data)[9]<-"Initial_Denial"
colnames(H1B_Data)[10]<-"Continuing_Approval"
colnames(H1B_Data)[11]<-"Continuing_Denial"

#Limit to 2022
H1B_Data<-subset(H1B_Data, Year <= 2022)
#Create Approval ratio columns
H1B_Data <- cbind(H1B_Data, H1B_Data$Initial_Approval/(H1B_Data$Initial_Approval+H1B_Data$Initial_Denial+H1B_Data$Continuing_Approval+H1B_Data$Continuing_Denial),  H1B_Data$Continuing_Approval/(H1B_Data$Initial_Approval+H1B_Data$Initial_Denial+H1B_Data$Continuing_Approval+H1B_Data$Continuing_Denial),  (H1B_Data$Initial_Approval+H1B_Data$Continuing_Approval)/(H1B_Data$Initial_Approval+H1B_Data$Initial_Denial+H1B_Data$Continuing_Approval+H1B_Data$Continuing_Denial))
#Rename columns accordingly
colnames(H1B_Data)[12]<-"IAppovals ratio"
colnames(H1B_Data)[13]<-"CApproval ratio"
colnames(H1B_Data)[14]<-"Approvals ratio"

#Create new dataset with industry
H1B_Industry<-H1B_Data[!is.na(H1B_Data$`Industry (NAICS) Code`),]
H1B_Industry$Professional <- with(H1B_Industry, ifelse(`Industry (NAICS) Code`> "54 - Professional, Scientific, and Technical Services", 1, 0))
industry_table<-table(H1B_Industry$`Industry (NAICS) Code`)
pie(industry_table)

H1B_numbers<-H1B_Data[c(1,8,9,10,11,12,13,14)]
H1B_numbers<-na.omit(H1B_numbers)

mean_ratios<-aggregate(H1B_numbers[6:8],by=list(H1B_numbers$Year), FUN=mean)
H1B_agg<-aggregate(H1B_numbers[2:8],by=list(H1B_numbers$Year), FUN=sum)

H1B_agg[6:8]<-mean_ratios[2:4]
colnames(H1B_agg)[1]<-"Year"

tsApprovals<-ts(H1B_agg$Initial_Approval + H1B_agg$Continuing_Approval, start=2009,end=2022)
tsDenials<-ts(H1B_agg$Initial_Denial + H1B_agg$Continuing_Denial, start=2009,end=2022)
tsApplications<-ts(H1B_agg$Initial_Approval + H1B_agg$Continuing_Approval + H1B_agg$Initial_Denial + H1B_agg$Continuing_Denial, start=2009,end=2022)
tsApproval_ratio<-ts(H1B_agg$`Approvals ratio`, start=2009,end=2022)
```
## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot


## Slide with R Output
```{r Viewing Data}
View(H1B_Industry)
View(H1B_numbers)
View(H1B_agg)
summary(H1B_agg)
H1B_agg %>%
  summary() %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
## Slide with Plot
```{r EDA H-1B}
par(mfrow=c(1,1))

plot.ts(tsApprovals)
plot.ts(tsDenials)
plot.ts(tsApplications)
plot.ts(tsApproval_ratio)
```

```{r Create Models}
tsApprovals_model<-tslm(tsApprovals~trend)
tsApplications_model<-tslm(tsApplications~trend)
tsRatio_model<-tslm(tsApproval_ratio~trend)

Ratio_model<-lm(`Approvals ratio`~Professional, H1B_Industry)

stargazer(Ratio_model,covariate.labels=c("Professional"), type="text",  out="models1.text")
summary(tsApprovals_model)
summary(tsApplications_model)
summary(tsRatio_model)

hist(H1B_Industry$Professional, breaks=2)

```

```{r Clean & Prepare LPR}
#Prepare LPR_data by renaming columns, subsetting, and changing data type
colnames(LPR_data)[2]<-"#_of_LPR_obtained"

tsLPR<-ts(LPR_data$`#_of_LPR_obtained`,start=2013,end=2022, frequency=1)
plot.ts(tsLPR)

LPR_tsmodel1<-tslm(tsLPR~trend)
summary(LPR_tsmodel1)

LPR_tsmodel2<-tslm(log(tsLPR)~trend)
summary(LPR_tsmodel2)

# forecast1<-forecast(LPR_tsmodel1, h=10)
# forecast2<-forecast(LPR_tsmodel2, h=10)
# View(forecast1)
# View(forecast2)
# 
# exp(forecast2$mean+se^2/2)

ses(LPR_data$`#_of_LPR_obtained`, alpha = 0.3)$fitted

fittedLPR1<-fitted(LPR_tsmodel1)
accuracy(fitted(LPR_tsmodel1),LPR_data$`#_of_LPR_obtained`[1:10])
se<-sigma(LPR_tsmodel2)
accuracy(fitted(LPR_tsmodel2)+se^2,LPR_data$`#_of_LPR_obtained`[1:10])
```

