---
title: "Winning the Lottery (BUS124B)"
author: "Mario Arce Acosta"
date: "2024-06-14"
output:
    slidy_presentation:
      theme: united
      highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Research Questions

- (Primary) Are certain industries associated with higher H-1B visa petition approval ratios? 
- (Relevance) How can the findings influence nonprofits for immigration, like the California Immigrant Policy Center, in their decision-making?
- (Secondary) What are some trends in immigration regarding H-1B visas?

---

## Data Methodology

- I exported two datasets from the US Citizenship and Immigration Services (USCIS); one from their H-1B Data Hub webpage, the other dataset is yearbook of people who obtained Lawful Permanent Residency (LPR) status
- H-1B Data: I exported data from the USCIS H-1B Employer Data Hub by splitting exports into sections of a few years since the final dataset would include over 900,000 observations, but reduced to a little over 200,000 after looking at the variance of the data
- I simply changed some values to numeric ones in Excel, as well as copying and pasting the values across years since large exports were not supported






```{r H-1B_Data, echo=FALSE}
library(readxl)
library(readr)
library(ggplot2)
library(forecast)
library(stargazer)
library(kableExtra)

#USER XCITE: laptop"C:\Users\XCITE-admin\Documents\GitHub\Regular_R_Work\H1B_Data.xlsx"
#USER mario
H1B_Data <- read_csv("C:/Users/mario/Documents/GitHub/Regular_R_Work/H1B_Data.csv")
View(H1B_Data)

#LPR_data <- read_excel("C:/Users/mario/Documents/GitHub/Regular_R_Work/LPR_1820-2022_data.xlsx", sheet = "LPR obtainments")
#View(LPR_data)

# fearData<- read_excel("C:/Users/mario/Desktop/Datasets/Migration_Fear_EPU_Data.xlsx", sheet = "Sheet1")
# View(fearData)

# removalsData <- read_excel("C:/Users/mario/Desktop/Datasets/Alien_removal_returns_fy2019_data.xlsx", sheet = "(R) Aliens removed or returned")
# View(removalsData)
```

```{r  H1B_Data Clean & Prepare, echo=FALSE}
##Work began at XCITE

#Name Columns
colnames(H1B_Data)[1]<-"Year"
colnames(H1B_Data)[8]<-"Initial_Approval"
colnames(H1B_Data)[9]<-"Initial_Denial"
colnames(H1B_Data)[10]<-"Continuing_Approval"
colnames(H1B_Data)[11]<-"Continuing_Denial"

#Create Approval ratio columns
H1B_Data <- cbind(H1B_Data, H1B_Data$Initial_Approval/(H1B_Data$Initial_Approval+H1B_Data$Initial_Denial+H1B_Data$Continuing_Approval+H1B_Data$Continuing_Denial),  H1B_Data$Continuing_Approval/(H1B_Data$Initial_Approval+H1B_Data$Initial_Denial+H1B_Data$Continuing_Approval+H1B_Data$Continuing_Denial),  (H1B_Data$Initial_Approval+H1B_Data$Continuing_Approval)/(H1B_Data$Initial_Approval+H1B_Data$Initial_Denial+H1B_Data$Continuing_Approval+H1B_Data$Continuing_Denial))
#Rename columns accordingly
colnames(H1B_Data)[12]<-"IAppovals ratio"
colnames(H1B_Data)[13]<-"CApproval ratio"
colnames(H1B_Data)[14]<-"Approvals ratio"

#Create new dataset with industry
H1B_Industry<-H1B_Data[!is.na(H1B_Data$`Industry (NAICS) Code`),]
H1B_Industry<-subset(H1B_Industry, Year==2015)
H1B_Industry<-H1B_Industry[!is.na(H1B_Industry$Continuing_Denial),]
H1B_Industry<-H1B_Industry[!is.na(H1B_Industry$`Tax ID`),]
H1B_Industry<-H1B_Industry[c(-2,-3,-5,-6,-7)]
H1B_Industry$Professional <- with(H1B_Industry, ifelse(`Industry (NAICS) Code`> "54 - Professional, Scientific, and Technical Services", 1, 0))
H1B_Industry$Health <- with(H1B_Industry, ifelse(`Industry (NAICS) Code`> "62 - Health Care and Social Assitance", 1, 0))
H1B_Industry$Finance <- with(H1B_Industry, ifelse(`Industry (NAICS) Code`> "52 - Finance and Insurance", 1, 0))
H1B_Industry$Construction <- with(H1B_Industry, ifelse(`Industry (NAICS) Code`> "23 - Construction", 1, 0))
H1B_Industry$Education <- with(H1B_Industry, ifelse(`Industry (NAICS) Code`> "61 - Educational Services", 1, 0))
H1B_Industry$Manufacturing <- with(H1B_Industry, ifelse(`Industry (NAICS) Code`> "31-33 - Manufacturing", 1, 0))
View(H1B_Industry)
industry_table<-table(H1B_Industry$`Industry (NAICS) Code`)
pie(industry_table)

H1B_numbers<-H1B_Data[c(1,8,9,10,11,12,13,14)]
H1B_numbers<-na.omit(H1B_numbers)

mean_ratios<-aggregate(H1B_numbers[6:8],by=list(H1B_numbers$Year), FUN=mean)
H1B_agg<-aggregate(H1B_numbers[2:8],by=list(H1B_numbers$Year), FUN=sum)

H1B_agg[6:8]<-mean_ratios[2:4]
colnames(H1B_agg)[1]<-"Year"

tsApprovals<-ts(H1B_agg$Initial_Approval + H1B_agg$Continuing_Approval, start=2009,end=2024)
tsDenials<-ts(H1B_agg$Initial_Denial + H1B_agg$Continuing_Denial, start=2009,end=2024)
tsApplications<-ts(H1B_agg$Initial_Approval + H1B_agg$Continuing_Approval + H1B_agg$Initial_Denial + H1B_agg$Continuing_Denial, start=2009,end=2024)
tsApproval_ratio<-ts(H1B_agg$`Approvals ratio`, start=2009,end=2024)
```

## H-1B (aggregated by year) Summary Table
```{r Viewing H-1B by year, aggregated}
H1B_agg %>%
  summary() %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## H-1B (Panel Data) Summary Table
```{r Viewing H-1B by Industry }
H1B_Industry %>%
  summary() %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Time Series Plots of H-1B petitions (2009-2024)
```{r EDA H-1B}
par(mfrow=c(2,2))

plot.ts(tsApprovals, xlab="Year", ylab="# of Approvals")
plot.ts(tsDenials, xlab="Year", ylab="# of Denials")
plot.ts(tsApplications, xlab="Year", ylab="# of Petitions/Applications")
plot.ts(tsApproval_ratio, xlab="Year", ylab="Approval Ratio")
```

## Model Creation and Residual Plots for H-1B data
```{r Create Models, echo=TRUE}
Professional_model<-lm(`Approvals ratio`~ Professional + Health + Finance + Manufacturing + Education, H1B_Industry)
```

## Model summary statistics

```{r H-1B model}
stargazer(Professional_model,covariate.labels=c("Professional", "Health", "Finance", "Manufacturing", "Education", "Construction"), type="text",  out="models1.text")

```
## Conclusions

- With such a low adjusted R squared value, we cannot conclude anything except that the variables we chose do not adequately explain all the variance for approval ratio for H-1B visa petitions.
- Some future work that could be done is expanding the model to account for variance across years

## Data source link
- https://www.uscis.gov/tools/reports-and-studies/h-1b-employer-data-hub

